T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases

### 2024-11-30 00_JCR-submission-folder_SECOND-SUBMISSION/00_T-reX_manuscript.tex
commit\

\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage{listings}
\usepackage{xcolor}
\definecolor{blue}{RGB}{33,150,209}
\definecolor{red}{RGB}{255,0,0}
\usepackage{xurl}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{lineno}
\usepackage{float}

\setlength{\linenumbersep}{5pt}
% Save the old \texttt command
\let\oldtexttt\texttt
% Redefine the \UrlFont to use a smaller typewriter font
\renewcommand{\UrlFont}{\ttfamily\small}
% Redefine the \texttt command to use a smaller font size and allow line breaks
\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
\usepackage{hyperref}

@@

% Short author
\shortauthors{S.C. McDowall et~al.}

@@

% Authors
\author[1]{Stewart Charles McDowall}[orcid=0000-0001-0000-0000]
\ead{s.c.mcdowall@cml.leidenuniv.nl}
\cormark[1]
\cortext[cor1]{Corresponding author:}
\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.}

\author[1]{Elizabeth Lanphear}
\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing.}

\author[1]{Stefano Cucurachi}
\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}

\author[1]{Carlos Felipe Blanco}
\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.}

@@

\begin{figure}
	\centering
	\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format. Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
	\includegraphics[width=0.9\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
\end{figure}

@@

%! CO2
The inclusion of carbon capture and storage (CCS) in the prospective databases
is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase
in the production of carbon dioxide `waste' over the period from 2020--2040
that is not seen in the baseline scenario. This result highlights the fact that
the (often) downward trends in global warming impacts calculated with
prospective databases using standard LCIA methods are dependent on the
assumptions made about the introduction of CCS technology. The actual
deployment of these technologies was only around 37~Mt CO$_2$/yr as of
2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of
the RCP scenarios~\citep{sacchi2023premisedocs}. The authors consider that the
over-representation of CCS and the under-representation of waste processing
technologies in the prospective databases is one significant limitation to the
validity of prospective LCA databases. Significant work is needed to improve
the representation of these technologies in the databases to ensure that the
results are more realistic and useful for decision-making.
%!!!!

\begin{figure*}[]
	\centering
	\includegraphics[width=1.8\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only). The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
\end{figure*}
%!!!!

@@

A comparison of the results for `Coal (black)' demand using the T-reX method
with the LCIA method `EDIP 2003 - coal no LT' is shown
in~\autoref{fig:comparison_methods}. In this case, both the trends and the
magnitude of the scores were very similar, both demonstrating a general
decrease in the use of coking coal in the battery's footprints. Such
comparability was also observed for other fossil-fuel-related demands (e.g.,
natural gas and petroleum), and to a lesser extent, for some of the metal
demand methods (e.g., zinc and cobalt). Correlation between standard LCIA
methods and T-reX methods is not generally to be expected, however, due to
fundamental differences in the way that the methods are constructed. In
standard LCIA methods, impact scores are derived exclusively from the magnitude
of the exchanges between the biosphere and the technosphere, that is,
extraction and emission. In the T-reX methods, the footprint scores are based
on an accounting of either waste generated or material demand, both of which
are technosphere-technosphere exchanges in terms of LCA modelling. For the
material demand methods especially, this distinction is critical. For example,
the application of the EDIP 2003 or the CSP methods for a given metal will
provide a score that is proportional to the amount extracted by mining, whereas
the T-reX method provides an aggregation of the exchanges with the market for
that metal. The T-reX method, therefore, considers cases of co-production,
recycling, and substitution, providing a picture of the supply chain pressures
that are not captured by the standard LCIA methods. This makes the T-reX
methods more sensitive to the modeling choices (e.g., economic allocation) that
are generally embedded in LCA databases.

\begin{figure*}
	\centering
	\includegraphics[width=1.8\columnwidth]{figures/T-reX-coalANDedip.pdf}
	\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
\end{figure*}

@@

\begin{table}
	\centering
	\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion').}\label{tab:results-case_study-comparison}
	\begin{tabular}{lll}
		\toprule
		\textbf{Indicator}         & \textbf{T-reX} & \textbf{Laurenti et al.} \\
		\midrule
		Total solid waste (kg/kg)  & 50.9           & 62.5                     \\
		Total liquid waste (kg/kg) & 3.53           & 3.63                     \\
		Hazardous waste (kg/kg)    & 1.47           & 62.6                     \\
		Recycled waste (kg/kg)     & 1.59           & 1.98                     \\
		\bottomrule
	\end{tabular}
\end{table}

>>>>

\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage{listings}
\usepackage{xcolor}
\definecolor{blue}{RGB}{33,150,209}
\definecolor{red}{RGB}{255,0,0}% \usepackage{listings}
\usepackage{xurl}
% \usepackage{pdfpages}
\usepackage{enumitem}
% \usepackage{lineno}
\usepackage{float}
% 
% \setlength{\linenumbersep}{5pt}
% Save the old \texttt command
\let\oldtexttt\texttt
% Redefine the \UrlFont to use a smaller typewriter font
\renewcommand{\UrlFont}{\ttfamily\small}
% Redefine the \texttt command to use a smaller font size and allow line breaks
\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}
\usepackage{hyperref}

@@

% Short author
\shortauthors{S2. McDowall et~al.}

@@

% Authors
\author[1]{Stewart Charles McDowall}[orcid=0000-0002-5688-1279]
\ead{s.c.mcdowall@cml.leidenuniv.nl}
\cormark[1]
\cortext[cor1]{Corresponding author:}
\credit{Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization}

\author[1]{Elizabeth Lanphear}[orcid=0009-0003-6364-0873]
\credit{Conceptualization, Methodology, Software, Validation, Writing - review \& editing}

\author[1]{Stefano Cucurachi}[orcid=0000-0003-2748-655X]
\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}

\author[1]{Carlos Felipe Blanco}[orcid=0000-0001-8199-8420]
\credit{Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition}

@@

\begin{figure*}
	\centering
	\caption{The workflow of T-reX presented in a generalised (a) and computational (b) format. Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
	\includegraphics[width=1.8\columnwidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
\end{figure*}

@@

%! CO2
The inclusion of carbon capture and storage (CCS) in the prospective databases
is apparent in~\autoref{fig:waste_totalANDcarbon} (b), with the rapid increase
in the production of carbon dioxide `waste' over the period from 2020--2040
that is not seen in the baseline scenario. This result highlights the fact that
the (often) downward trends in global warming impacts calculated with
prospective databases using standard LCIA methods are dependent on the
assumptions made about the introduction of CCS technology. The actual
deployment of these technologies was only around 37~Mt CO\(_2\)/yr as of
2023~\citep{dziejarski2023ccs}, far short of the levels projected in many of
the RCP scenarios~\citep{sacchi2023premisedocs}. The authors consider that the
over-representation of CCS and the under-representation of waste processing
technologies in the prospective databases is one significant limitation to the
validity of prospective LCA databases. Significant work is needed to improve
the representation of these technologies in the databases to ensure that the
results are more realistic and useful for decision-making.
%!!!!

\begin{figure*}[]
	\centering
	\includegraphics[width=2.1\columnwidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
	\caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (produced from carbon capture and storage (CCS) only). The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
\end{figure*}
%!!!!

@@

A comparison of the results for `Coal (black)' demand using the T-reX method
with the LCIA method `EDIP 2003 – coal no LT' is shown
in~\autoref{fig:comparison_methods}. In this case, both the trends and the
magnitude of the scores were very similar, both demonstrating a general
decrease in the use of coking coal in the battery's footprints. Such
comparability was also observed for other fossil-fuel-related demands (e.g.,
natural gas and petroleum), and to a lesser extent, for some of the metal
demand methods (e.g., zinc and cobalt). Correlation between standard LCIA
methods and T-reX methods is not generally to be expected, however, due to
fundamental differences in the way that the methods are constructed. In
standard LCIA methods, impact scores are derived exclusively from the magnitude
of the exchanges between the biosphere and the technosphere, that is,
extraction and emission. In the T-reX methods, the footprint scores are based
on an accounting of either waste generated or material demand, both of which
are technosphere-technosphere exchanges in terms of LCA modelling. For the
material demand methods especially, this distinction is critical. For example,
the application of the EDIP 2003 or the CSP methods for a given metal will
provide a score that is proportional to the amount extracted by mining, whereas
the T-reX method provides an aggregation of the exchanges with the market for
that metal. The T-reX method, therefore, considers cases of co-production,
recycling, and substitution, providing a picture of the supply chain pressures
that are not captured by the standard LCIA methods. This makes the T-reX
methods more sensitive to the modeling choices (e.g., economic allocation) that
are generally embedded in LCA databases.

\begin{figure*}
	\centering
	\includegraphics[width=2.0\columnwidth]{figures/T-reX-coalANDedip.pdf}
	\caption{Comparison of the (a) `T-reX  	\caption{Comparison of the (a) `T-reX  Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
\end{figure*}

@@

\begin{table}
	\centering
	\caption{Comparison of the results from the T-reX battery case study (database: \texttt{ecoinvent} 3.9.1 REMIND SSP2 Base 2020, activity: Li-ion NMC 111) with those from \cite{laurenti2023wastefootprint} (database: \texttt{ecoinvent} 3.5, activity:`Li-ion')}\label{tab:results-case_study-comparison}
	\begin{tabular}{lll}
		\toprule
		\textbf{Indicator}         & \textbf{T-reX} & \textbf{Laurenti et al.} \\
		\midrule
		Total solid waste (kg/kg)  & 50.9           & 62.5                     \\
		Total liquid waste (kg/kg) & 3.53           & 3.63                     \\
		Hazardous waste (kg/kg)    & 1.47           & 62.6                     \\
		Recycled waste (kg/kg)     & 1.59           & 1.98                     \\
		\bottomrule
	\end{tabular}
\end{table}
